### 2019-01-25 metrics2.tex
metrics

\begin{enumerate}[(a)]
	\item Denote sample variances as $s^2_{ab}, a, b \in \left\{y, x_1, x_2\right\}$. To compute the OLS estimate we need to solve the following optimization problem:
	\begin{align*}
	\underset{b_0, b_1, b_2}{\min} \summa (y_i - b_0 - b_1x_{1i} - b_2x_{2i})^2\\
	\text{FOCs:} \begin{cases}
	\summa (y_i - b_0 - b_1x_{1i} - b_2x_{2i}) = 0\\
	\summa (y_i - b_0 - b_1x_{1i} - b_2x_{2i})x_{1i} = 0\\
	\summa (y_i - b_0 - b_1x_{1i} - b_2x_{2i})x_{2i} = 0
	\end{cases}\\
	b_0 = \bar{y} - b_1\bar{x_1} - b_2 \bar{x_2}\\
	b_1\left(\summa x_{1i}^2\right) = \summa (y_i - \bar{y} + b_1\bar{x_1} + b_2\bar{x_2} - b_2x_{2i})x_{1i}
	\end{align*}
	\begin{align*}
	b_1 = \frac{\summa (y_i - \bar{y} - b_2(x_{2i} - \bar{x}))x_{1i}}{\summa (x_{1i} - \bar{x})^2}&=\frac{\summa (y_i - \bar{y})(x_{1i} - \bar{x_1}) - b_2\summa(x_{2i} - \bar{x_2})(x_{1i} - \bar{x_1})}{\summa (x_{1i} - \bar{x})^2} =\\
	&=\frac{s^2_{yx_1} - b_2s^2_{x_1x_2}}{s^2_{x_1x_1}}
	\end{align*}
	Similarly for $b_2$:
	\begin{align}\label{eq2}
	b_2 &= \frac{s^2_{yx_2} - b_1s^2_{x_1x_2}}{s^2_{x_2x_2}}\nonumber\\
	\text{hence }\nonumber\\
	b_2 &= \frac{s^2_{yx_2}s^2_{x_1x_1} - s^2_{yx_1}s^2_{x_2x_1}}{s^2_{x_1x_1}s^2_{x_2x_2} - (s^2_{x_1x_2})^2}
	\end{align}
	\item Performing simple linear regression of $y$ on $x_1$, $y$ on $x_2$, $x_1$ on $x_2$ and $x_2$ on $x_1$ denote OLS estimates for slope coefficients as $\alpha_1, \alpha_2, \alpha_3, \alpha_4$ respectively. It is known that
	\begin{align*}
	\alpha_1 = \frac{s^2_{yx_1}}{s^2_{x_1x_1}},\ \alpha_2 = \frac{s^2_{yx_2}}{s^2_{x_2x_2}},\ \alpha_3 = \frac{s^2_{x_1x_2}}{s^2_{x_2x_2}},\ \alpha_4 = \frac{s^2_{x_1x_2}}{s^2_{x_1x_1}}
	\end{align*}
	Divide numerator and denumerator of (\ref{eq2}) by $(s^2_{x_1x_1})^2$ (assuming that it is not equal to 0):
	\begin{align*}
	b_2 = \frac{\alpha_2\frac{\alpha_4}{\alpha_3} - \alpha_1\alpha_4}{\frac{\alpha_4}{\alpha_3} - \alpha_4^2} = \frac{\frac{\alpha_2}{\alpha_3} - \alpha_1}{\frac{1}{\alpha_3} - \alpha_4}
	\end{align*}
	Similarly for $b_1$:
	\begin{align*}
	b_1 = \frac{\alpha_1 - \alpha_2 \alpha_4}{1 - \alpha_3}
	\end{align*}
	\item From (\ref{eq2}):
	\begin{align*}
	\hat{\gamma_2} = \frac{s^2_{yx_2}s^2_{x_1x_1} - s^2_{x_2x_1}s^2_{yx_1}}{s^2_{x_1x_1}s^2_{yy} - (s^2_{x_1y})^2} \to \frac{1}{\hat{\gamma_2}} = \frac{s^2_{x_1x_1}s^2_{yy} - (s^2_{x_1y})^2}{s^2_{yx_2}s^2_{x_1x_1} - s^2_{x_2x_1}s^2_{yx_1}}\\
	\end{align*}
	To find $\expect[\frac{1}{\hat{\gamma_2}}]$ firstly observe the following facts:
	\begin{align*}
	s^2_{x_1x_1}s^2_{yy} = s^2_{x_1x_1}\cdot(\summa (y_i - \bar{y})(\beta_1x_{1i} + \beta_2x_{2i} + u_i - \beta_1\bar{x_1} - \beta_2 \bar{x_2} - \bar{u})) = \\
	=s^2_{x_1x_1} (\beta_1s^2_{yx_1} + \beta_2s^2_{yx_2} + s^2_{uy})\\
	(s_{x_1y}^2)^2 = s_{x_1y}^2 \cdot (\summa (x_{1i} - \bar{x})(\beta_1x_{1i} + \beta_2x_{2i} + u_i - \beta_1\bar{x_1} - \beta_2\bar{x_2} - \bar{u})) =\\ =s_{x_1y}^2 (\beta_1 s^2_{x_1x_1} + \beta_2s^2_{x_1x_2} + s^2_{x_1u})
	\end{align*}
	hence
	\begin{align*}
	s^2_{x_1x_1}s^2_{yy} - (s_{x_1y}^2)^2 = \beta_2(s^2_{x_1x_1}s^2_{yx_2} - s^2_{x_1y}s^2_{x_1x_2}) + s^2_{x_1x_1}s^2_{uy} - s^2_{x_1y}s^2_{x_1u}
	\end{align*}
	Thus,
	\begin{align*}
	\frac{1}{\hat{\gamma_2}} = \beta_2 + \frac{s^2_{x_1x_1}s^2_{uy} - s^2_{x_1y}s^2_{x_1u}}{s^2_{x_1x_1}s^2_{yx_2} - s^2_{x_1y}s^2_{x_1x_2}}
	\end{align*}
\end{enumerate}
\section*{Problem 4}
Consider a simple regression model through the origin under the Gauss-Markov assumptions
\begin{align*}
y = \gamma x + \varepsilon
\end{align*}
and let
\begin{align*}
\tilde{\gamma} = \frac{\summa y_ix_i^3}{\summa x_i^4}
\end{align*}
be an estimator of the slope parameter $\gamma$.
\begin{enumerate}[(a)]
\item Under what condition this estimator is well defined?
\item Show conditional unbiasness of $\tilde{\gamma}$.
\item Derive conditional variance of the estimator.
\end{enumerate}

@@

\begin{enumerate}[(a)]
	\item Coefficient on \yrseduc\ is equal to 1.34147. That means that holding a gender of the person fixed each additional year of education on average gives 1.34147 dollars of increase of the per hour earnings.
	\item Using coeftest() it can be obtained that the null hypothesis should be rejected on the 5\% significance level. The hypothesis was that holding a number of years of education fixed both males and females get on average equal salary per hour.
	\item Coefficient on \yrseduc\ did not significantly change: from 1.32 to 1.34. The reason is the follwoing that as it has been already indicated in the problem 5 (a) the sample correlation between \yrseduc\ and \textit{female} is very low, as a result, using formula from the 5 (c) $\tilde{\beta_1} = \hat{\beta_1} + \hat{\beta_2} \delta$ one can conclude that extremely small $\delta$ results in $\tilde{\beta_1} \approx \hat{\beta_1}$ even if $\hat{\beta_2} \neq 0$.
\end{enumerate}
\end{document}

>>>>

\begin{enumerate}[(a)]
	\item Denote sample variances as $s^2_{ab}, a, b \in \left\{y, x_1, x_2\right\}$. To compute the OLS estimate we need to solve the following optimization problem:
	\begin{align*}
	\underset{b_0, b_1, b_2}{\min} \summa (y_i - b_0 - b_1x_{1i} - b_2x_{2i})^2\\
	\text{FOCs:} \begin{cases}
	\summa (y_i - b_0 - b_1x_{1i} - b_2x_{2i}) = 0\\
	\summa (y_i - b_0 - b_1x_{1i} - b_2x_{2i})x_{1i} = 0\\
	\summa (y_i - b_0 - b_1x_{1i} - b_2x_{2i})x_{2i} = 0
	\end{cases}\\
	b_0 = \bar{y} - b_1\bar{x_1} - b_2 \bar{x_2}\\
	b_1\left(\summa x_{1i}^2\right) = \summa (y_i - \bar{y} + b_1\bar{x_1} + b_2\bar{x_2} - b_2x_{2i})x_{1i}
	\end{align*}
	\begin{align*}
	b_1 = \frac{\summa (y_i - \bar{y} - b_2(x_{2i} - \bar{x}))x_{1i}}{\summa (x_{1i} - \bar{x})^2}&=\frac{\summa (y_i - \bar{y})(x_{1i} - \bar{x_1}) - b_2\summa(x_{2i} - \bar{x_2})(x_{1i} - \bar{x_1})}{\summa (x_{1i} - \bar{x})^2} =\\
	&=\frac{s^2_{yx_1} - b_2s^2_{x_1x_2}}{s^2_{x_1x_1}}
	\end{align*}
	Similarly for $b_2$:
	\begin{align}\label{eq2}
	b_2 &= \frac{s^2_{yx_2} - b_1s^2_{x_1x_2}}{s^2_{x_2x_2}}\nonumber\\
	\text{hence }\nonumber\\
	b_2 &= \frac{s^2_{yx_2}s^2_{x_1x_1} - s^2_{yx_1}s^2_{x_2x_1}}{s^2_{x_1x_1}s^2_{x_2x_2} - (s^2_{x_1x_2})^2}
	\end{align}
	\item Performing simple linear regression of $y$ on $x_1$, $y$ on $x_2$, $x_1$ on $x_2$ and $x_2$ on $x_1$ denote OLS estimates for slope coefficients as $\alpha_1, \alpha_2, \alpha_3, \alpha_4$ respectively. It is known that
	\begin{align*}
	\alpha_1 = \frac{s^2_{yx_1}}{s^2_{x_1x_1}},\ \alpha_2 = \frac{s^2_{yx_2}}{s^2_{x_2x_2}},\ \alpha_3 = \frac{s^2_{x_1x_2}}{s^2_{x_2x_2}},\ \alpha_4 = \frac{s^2_{x_1x_2}}{s^2_{x_1x_1}}
	\end{align*}
	Divide numerator and denumerator of (\ref{eq2}) by $(s^2_{x_1x_1})^2$ (assuming that it is not equal to 0):
	\begin{align*}
	b_2 = \frac{\alpha_2\frac{\alpha_4}{\alpha_3} - \alpha_1\alpha_4}{\frac{\alpha_4}{\alpha_3} - \alpha_4^2} = \frac{\frac{\alpha_2}{\alpha_3} - \alpha_1}{\frac{1}{\alpha_3} - \alpha_4}
	\end{align*}
	Similarly for $b_1$:
	\begin{align*}
	b_1 = \frac{\alpha_1 - \alpha_2 \alpha_4}{1 - \alpha_3}
	\end{align*}
	\item 
	\begin{align*}
	\beta_2 = \frac{1}{\gamma_2}\\
	\expect[\hat{\beta_2}] = \beta_2\\
	\expect[\hat{\gamma_2}] = \gamma_2\\
	\beta_2 = \frac{1}{\expect[\hat{\gamma_2}]}
	\end{align*}
	Since $\frac{1}{x}$ is a strictly concave function for $x < 0$ and strictly convex for $x > 0$ and moreover $\hat{\gamma_2}$ is not almost sure constant (we assume it) then by Jensen's inequality:
	\begin{align*}
	\expect\left[\frac{1}{\hat{\gamma_2}}\right] < \frac{1}{\expect[\hat{\gamma_2}]}, \hat{\gamma_2} < 0\\
	\expect\left[\frac{1}{\hat{\gamma_2}}\right] > \frac{1}{\expect[\hat{\gamma_2}]}, \hat{\gamma_2} > 0\\
	\end{align*}
	Thus, 
	\begin{align*}
	\expect\left[\frac{1}{\hat{\gamma_2}}\right] < \beta_2, \hat{\gamma_2} < 0\\
	\expect\left[\frac{1}{\hat{\gamma_2}}\right] > \beta_2, \hat{\gamma_2} > 0\\
	\end{align*}
	hence it is biased.
\end{enumerate}
\section*{Problem 4}
Consider a simple regression model through the origin under the Gauss-Markov assumptions
\begin{align*}
y = \gamma x + \varepsilon
\end{align*}
and let
\begin{align*}
\tilde{\gamma} = \frac{\summa y_ix_i^3}{\summa x_i^4}
\end{align*}
be an estimator of the slope parameter $\gamma$.
\begin{enumerate}[(a)]
\item Under what condition this estimator is well defined?
\item Show conditional unbiasness of $\tilde{\gamma}$.
\item Derive conditional variance of the estimator.
\end{enumerate}

@@

\begin{enumerate}[(a)]
	\item Coefficient on \yrseduc\ is equal to 1.34147. That means that holding a gender of the person fixed each additional year of education on average gives 1.34147 dollars of increase of the per hour earnings.
	\item Using coeftest() it can be obtained that the null hypothesis should be rejected on the 5\% significance level. The hypothesis was that holding a number of years of education fixed both males and females get on average equal salary per hour.
	\item Coefficient on \yrseduc\ did not significantly change: from 1.32 to 1.34. The reason is the follwoing that as it has been already indicated in the problem 5 (a) the sample correlation between \yrseduc\ and \textit{female} is very low, as a result, using formula from the 5 (c) $\tilde{\beta_1} = \hat{\beta_1} + \hat{\beta_2} \delta$ one can conclude that extremely small $\delta$ results in $\tilde{\beta_1} \approx \hat{\beta_1}$ even if $\hat{\beta_2} \neq 0$.
\end{enumerate}
\section*{Problem 8}
Use multiple linear regression of \textit{ahe} on \yrseduc\ and \textit{ba} (a binary variable equaling one for
those who attained a bachelor’s degree) to estimate the marginal value of a bachelor’s degree,
more specifically, the value of going to school for 16 years and receiving a bachelor’s degree,
relative to going to school for 16 years but not receiving a bachelor’s degree. Construct a 95\%
confidence interval for the marginal value of a bachelor’s degree.
\end{document}
